Initial Analysis

### 2022-06-30 Scripts/Initial_Analysis.Rmd
Incorporated straddling analysis to transboundary analysis

@@

GetTransResults=function(Spp){

  # Set the path for each file
  Distpath <- paste(Results_Path,"Trans_Results/",Spp,sep="")

  # Loads all files in a df
  Load_Data <- bind_rows(lapply(Distpath, FUN=fread))

  return(Load_Data)

@@

````
## Load Results data

@@

@@

@@

Species_Sampled <- fread(paste(my_path("D"),"Distribution/Sampled_Species.csv", sep =""))

@@

Clean_Results_Trans <- Raw_Results_Trans %>% 
  CleanResults(.,matching_names) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(matching_names,
            by = "sf_eez_name") %>% 
  rename(eez_name = sf_eez_name,
         fishing_entity = sau_fishing_entity)

@@

SAU_Global_Data <-fread(paste(my_path("D"),"Distribution/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
  # Remove high seas and 207 and 208 that are unknown
  filter(!fishing_entity_id %in% c(213,223)) %>%  # 207,208
  # Allocate catch of islands to UK
  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
  # Convert to 2019 real USD
  mutate(values = 1.16 * value) %>%
  # Average 10 years of data
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("Variable","Value",catch:value)

@@

@@

>>>>

## Methods for straddling stocks

1.- Get transboundary species list

1.1- Remove highly migratory species based on literature

2- High seas shapefile

2.1- FAO Areas ID

3.- Identiy what transboundary species are actually straddling

4.- Be happy, happy like a hippo

### Shapefiles

```{r}

# World land maps
world_land <- ne_countries(scale = 'medium', returnclass = c("sf")) %>% 
  st_transform(crs = 4326)# 4326

# ggplot(world_land) +
#   geom_sf()

# FAO regions
fao_regions <- my_sf("FAO") %>% 
  filter(F_LEVEL == "MAJOR")
st_crs(fao_regions) = 4326

# head(fao_regions)

# fao_regions %>% 
#   ggplot() +
#   geom_sf(aes(fill = F_AREA))

# EEZs
world_eez <- my_sf("SAU",simple = 1000)

# Lon-lat DBEM

# First get Vicky's data to get high seas
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE)
colnames(EEZ_CellID) <- c("EEZID","index")

# Now our gridcell
abny <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F) %>% 
  bind_cols(my_path("Spa","DBEM","WArea.txt",read = T, header = F)) %>% 
  filter(!index %in% EEZ_CellID$INDEX) %>% 
  mutate(cat = ifelse(v1>0, "abnj","land")) %>% 
  select(-v1)

# SAVE FOR COMPLETE DBEM
complete_lon_lat_dbem <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F) %>% 
  left_join(abny) %>% 
  mutate(cat = ifelse(is.na(cat), "eez",cat))

# Save for fouture
# write_csv(complete_lon_lat_dbem, my_path("Spa", "DBEM","complete_lon_lat_dbem.csv"))

  ggplot(complete_lon_lat_dbem) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = cat
    )
  )

# All maps test
# ggplot() +
#   geom_sf(data = fao_regions, aes()) +
#   geom_sf(data = world_land, aes()) +
#   geom_sf(data = world_eez, aes())

  # First get Vicky's data to get high seas
EEZ_SAU <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = TRUE)
colnames(EEZ_SAU) <- c("sau_code","eez")

# First get Vicky's data to get high seas
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE)
colnames(EEZ_CellID) <- c("sau_code","index")

complete_sau <- EEZ_SAU %>% 
  left_join(EEZ_CellID)

#   
# coords %>% 
#   left_join(EEZ_CellID) %>% 
#     rename(sau_eez_id = EEZID) %>% 
#     gather("code_na","code",f_code:sau_eez_id) %>%

  # write_csv(Coord, my_path("Spa", "DBEM","complete_lon_lat_dbem.csv"))
```

### Determining straddling stocks

```{r GetSppDistFun}

  Distpath <- paste(Data_Path,"Data/Distribution/",sep="")
  INDEX = seq(1,259200,1)
  # INDEX <- Coordinates$INDEX

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- suppressWarnings(sppdist[order(sppdist$latitude, rev(sppdist$longitude),decreasing=TRUE), ] %>% 
                                    mutate(INDEX = seq(1,259200,1)) %>% 
                                    gather("Model","Value",3:6) %>%
                                    mutate(Value = ifelse(is.na(Value), 0, Value)) # Converting NA's to ceros
      )

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- sppdist[order(sppdist$latitude, rev(sppdist$longitude),decreasing=TRUE), ] %>% 
        mutate(INDEX = seq(1,259200,1))

  }
}
```

```{r StraddIndexFun}

StraddIndex <- function(Spp,Model = "All",Neighbours,Coord,Index_Code){

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(#INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
      Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = sum(Value,na.rm=T)/3
    ) %>%
    filter(Model_Index > 0.4) %>% # at least 2 sources agree
    # ____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>%
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    select(-Model, -Value, -latitude,-longitude) %>%
    rename(index = INDEX) %>% 
    # joint eez names
    left_join(Index_Code,
              by = "index") %>% 
    # joint fao area names
    left_join(fao_areas,
              by = "index") %>% 
    mutate(
      territory = ifelse(!is.na(name),name,name_en),
      category = ifelse(!is.na(name),"eez","fao"),
    ) %>% 
    select(1:3,territory,category) %>% 
    clean_names() %>% 
    filter(model_index == 100)

  # Distributional test
  # test <- Trans_Spp %>% 
    # left_join(coords)

  # ggplot(test) +
  #   geom_tile(
  #     aes(
  #       x = lon,
  #       y = lat,
  #       fill = category
  #     )
  #   )

  # Step 2.1. Determines not a straddling stock
  if("fao" %in% Trans_Spp$category == FALSE){

    # output <- Spp_Grid %>% 
    #   select(
    #     taxon_key,
    #     territory,
    #     neighbour
    #   ) %>% 
    #   mutate(straddling = "no")
    return(print("not a straddling stock"))
    stop()

    # MODEL INDEX (TRESHOLDS 1 & 2) #
    Model_Index_D <- Trans_Spp %>% 
      group_by(territory,
               category,
               taxon_key,
               model_index
      ) %>% 
      summarise(n_cells_spp = n()) %>% 
      select(-n_cells_spp)

    #Step 1.  Get EEZ id and Neighbour
    Neighbours_List <- neighbours %>%
      group_by(territory,neighbour) %>%
      summarise(n=n()) %>%
      ungroup() %>%
      select(-n) %>% 
      left_join(fao_to_sau) %>% 
      filter(!is.na(territory))

    # Set FAO neighbours... 
    fao_neighbours <- Neighbours_List %>% 
      select(territory,fao_area_name) %>% 
      rename(territory = fao_area_name,
             neighbour = territory) %>% 
      group_by(territory, neighbour) %>% 
      summarise(n())

    fao_territory <- fao_neighbours %>% 
      rename(territory = neighbour,
             neighbour = territory)

    neighbours_list_fao <- fao_neighbours %>% 
      bind_rows(Neighbours_List,fao_territory) %>% 
      select(-fao_area_name)

    # Step 2. Determines the amount of grids present in each country
    Spp_Grid <- Trans_Spp %>% 
      group_by(taxon_key,
               model_index, # Un-comment after producing models x datasets
               territory,
               category) %>% 
      summarise(n_spp_eez = length(unique(index))) %>% 
      left_join(neighbours_list_fao,
                by = "territory") %>% 
      filter(territory %in% Trans_Spp$territory, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
             neighbour %in% Trans_Spp$territory)

    # Split dataframes to merge latter
    Territory_T <- Spp_Grid %>% 
      filter(category == "eez") %>% 
      ungroup() %>% 
      select(
        model_index, # Un-comment after producing models x datasets
        taxon_key,
        Name=territory,
        n_spp_eez
      )

    Neighbour_T <- Spp_Grid %>%
      filter(category == "eez") %>% 
      ungroup() %>% 
      select(
        model_index,
        taxon_key,
        n_spp_eez,
        territory,
        Name=neighbour,
        territory
      )

    fao_grid <- Spp_Grid %>% 
      filter(category == "fao") %>% 
      left_join(neighbours_list_fao) %>% 
      select(Name =neighbour,
             fao_name = territory,
             n_spp_eez.z = n_spp_eez)

    # Merge dataframes to get totals per Neighbourds
    Area_Index_D <- full_join(Territory_T,
                              Neighbour_T, 
                              by = c("model_index","Name","taxon_key")
    ) %>%
      left_join(fao_grid) %>% 
      filter(!is.na(fao_name)) %>% # remove non-straddling stocks
      rowwise() %>%
      mutate(spp_total = sum(n_spp_eez.x,n_spp_eez.y,n_spp_eez.z,na.rm=T)) %>% # Total gridcelles per Neighbours
      distinct() %>% # Removes false duplicates from `full_join()`
      rename(territory = Name,
             neighbour = territory,
             n_spp_country = n_spp_eez.x,
             n_spp_neighbour = n_spp_eez.y,
             n_spp_fao_area = n_spp_eez.z) %>%
      # Estimate if it is straddling
      mutate(stradling_index = (n_spp_fao_area/spp_total)*100)

    output <- Area_Index_D %>% 
      select(taxon_key,
             territory,
             neighbour,
             fao_name,
             stradling_index)

    File_Name <- paste(Spp,"_straddling.csv",sep = "")
    Save_Path <- my_path("R","EEZ/Stradling_Results/", File_Name)

    # Save_Path <- paste(Results_Path,"Trans_Results/",File_Name,sep="")

    write_csv(output,
              Save_Path)
  }
} # closes function

 #Test me  
# suppressMessages(
#   StraddIndex(600069, Model = "All", Coord = coords, Index_Code = Index_Code)
# )

### Straddling Control pannel

```{r}
# DBEM coordinate system
# complete_lat_lon <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T)

# Gabriel's coordinate system
CoorG <- my_path("G","FishForVisa/Data/Spatial/","coordinates_gab.csv",read =T) %>%
  mutate(INDEX = seq(1,259200,1))

# Get fao areas on DBEM grid
fao_areas <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T) %>% 
  filter(cat == "abnj")

# Fao combo
fao_to_sau <- read_excel("/Volumes/DATA/JULIANO_NEYMAR/Spatial/SAU/EEZ-FAO-Country-Combo.xlsx") %>% 
  select(territory = eez_name,eez_id,fao_area_name)

# Neighboring dataset
neighbours <- read_excel("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Data/Spatial/EEZ_Neighbour_List.xlsx") %>% 
  clean_names() %>% 
  rename(eez_id = eezid) %>% 
  left_join(fao_to_sau,
            by = "eez_id") %>% 
  select(territory = territory.x,neighbour)

# Load FishForVisa Results
transboundary_spp <- read_csv("https://media.githubusercontent.com/media/jepa/FishForVisa/master/Data/Results/Clean_Results_Trans.csv") %>% 
  # Set same filters as in FishForVisa
  filter(
    area_index >= 0.25,
    area_index <= 1-0.25,
    model_index >= 100
  )

# DepPelingo

depth_profile <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

# Double check we have 633 species
length(unique(transboundary_spp$taxon_key)) # 633

coords <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F)

Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/" #x

# SAU relations between INDEX and Country's EEZs
# First get Vicky's data to get high seas
EEZIDs_List <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = T) #x
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE) #x
colnames(EEZ_CellID) <- c("eezid","index") #x

Index_Code <- EEZIDs_List %>%  #x
  left_join(EEZ_CellID)

```{r run_me, eval = T}

suppressWarnings(
  suppressMessages(
    lapply(unique(transboundary_spp$taxon_key), StraddIndex, Model = "All", Coord = coords,
           Index_Code = Index_Code)
  )
)

@@

GetTransResults=function(Spp, type = "NA"){

  if(type == "strad"){
    # Set the path for each file
  Distpath <- paste(my_path("R","EEZ/Straddling_Results/"),Spp,"_straddling.csv",sep="")

  }else{
  # Set the path for each file
  Distpath <- paste(Results_Path,"Trans_Results/",Spp,sep="")

  # Loads all files in a df
  Load_Data <- bind_rows(lapply(Distpath, FUN=fread))
  return(Load_Data)

@@

### High migratory species

Remove highly migratory species according to UNCLOS extracted from FAO (Maguire et all). The above tuna species listed as highly migratory species in Annex 1 of UNCLOS have extensive distribution on the high seas

Maguire, J-J., Sissenwine, M., Csirke, J., Grainger, R., & Garcia, S. (2006). The state of world highly migratory, straddling and other high seas fishery resources and associated species. FAO.

```{r highly_migratory_species}

high_mig_taxa <- c("Thunnus alalunga", "Thunnus thynnus","Thunnus orientalis","Thunnus obesus","Katsuwonus pelamis", "Thunnus albacares","Thunnus atlanticus","Euthynnus alletteratus","Euthynnus affinis","Thunnus maccoyii", "Auxis thazard", "Auxis rochei",# Tuna-likes
                   "Tetrapturus angustirostris", "Tetrapturus belone", "Tetrapturus pfluegeri", "Tetrapturus albidus", "Tetrapturus audax", "Tetrapturus georgei", "Makaira indica", "Makaira nigricans","Istiophorus platypterus", "Istiophorus albicans","Xiphias gladius")

length(high_mig_taxa) # 23

high_mig_spp <- my_path("D","Distribution/Species","SppTaxonName.csv", read = T) %>% 
  filter(taxon_name %in% high_mig_taxa)

nrow(high_mig_spp) # 21

# ---------------- 
# Missing 4 species
# missing_hm_spp <- as.data.frame(x = high_mig_taxa) %>% 
  # filter(!high_mig_taxa %in% high_mig_spp$taxon_name)

# Tetrapturus belone # only found in the med				
# Tetrapturus georgei # Misspselled, correct name Tetrapturus georgii but not identified in the SAU dataset

# ---------------- # missing data
```

## Load Results data

@@

# ----------------- #
# STRADDLING
# ----------------- #
fao_areas <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T) %>% 
  filter(cat == "abnj")

# Get species taxon data
taxon_list <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

# Load straddling species
# Spp list
spp_complete <- list.files(my_path("R","EEZ/Straddling_Results/"))

stradling_spp <- str_replace(spp_complete,"\\_.*","")

stradd_data <- bind_rows(
  lapply(stradling_spp,GetTransResults, type = "strad")
  ) %>% 
  left_join(taxon_list) %>% 
  select(taxon_key,taxon_name,dem_pel, everything()) %>% 
  mutate(category = ifelse(taxon_key %in% high_mig_spp$taxon_key,"high_migratory","straddling"))

# Complete transboundary data

stradd_hs_spp <- stradd_data %>% 
  select(taxon_key,category,sf_eez_name = territory, eez_neighbour = neighbour) %>% 
  unique()

@@

# Get species taxon data
taxon_list <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

@@

# STRADDLING: needed
Species_Sampled <- fread(paste(my_path("D"),"Distribution/Sampled_Species.csv", sep =""))

@@

# Clean_Results_Trans <- Raw_Results_Trans %>% 
#   CleanResults(.,matching_names) %>% 
#   rename(sf_eez_name = eez_name) %>% 
#   left_join(matching_names,
#             by = "sf_eez_name") %>% 
#   rename(eez_name = sf_eez_name,
#          fishing_entity = sau_fishing_entity)

Clean_Results_Trans <- my_path("R","EEZ","Clean_Results_Trans.csv", read = T) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(matching_names)

@@

# Used for STRADDLING too
SAU_Global_Data <-fread(paste(my_path("D"),"Distribution/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
  # Remove high seas and 207 and 208 that are unknown
  filter(!fishing_entity_id %in% c(213,223)) %>%  # 207,208
  # Allocate catch of islands to UK
  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
  # Convert to 2019 real USD
  mutate(values = 1.16 * value) %>%
  # Average 10 years of data
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("Variable","Value",catch:value)

@@

## STRADLING
shared_species <- Tresholded_T_Data %>% 
  left_join(stradd_hs_spp) %>% 
  mutate(category = replace_na(category,"transboundary"))

shared_classification <- shared_species %>% 
  select(taxon_key,category,sf_eez_name) %>% 
  unique()

@@

### Straddling Species

```{r}

# Number of species that are straddling or highly migratory
stradd_data %>% 
  group_by(category) %>% 
  summarize(
    n_spp = length(unique(taxon_key)),
    prop = round((length(unique(taxon_key))/633)*100)
    )

# RESULT
# category  n_spp   prop of transboundary
# high_mig	 20	     3		
# stradd	   84	    13	
# total     104     16

EEZIDs_List <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = T) %>% 
  rename(eez_name = name)

# Econ value
Summary_Trans <- SAU_Sampled_Data %>% 
  semi_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity_id")
  ) %>% 
  group_by(Variable) %>%
  summarise(Total_Trans = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Trans/Total_Sampled*100,
         Per_SAU = Total_Trans/Total_Global*100)
